---
title: "Hands-on Exercise 1 Introduction to Visual Analytics"
author: "Yen Yun Hsuan"
---

# Getting Started

## Install and launching R packages

The code chunk below uses p_load() of pacman package to check if tidyverse packages are installed in the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(tidyverse)
```

## Importing the data

```{r}
library(readr)
exam_data <- read_csv("data/Exam_data.csv")
```

# Plotting a simple bar chart

## Bar chart

```{r}
ggplot(data=exam_data,
       aes(x=RACE))+
  geom_bar()

```

### Horizontal bar chart

Use coordinate to change bar into horizontal

```{r}
ggplot(data=exam_data,
       aes(x=RACE))+
  geom_bar()+
  coord_flip()+
  theme_minimal()
```

## Histogram

```{r}
hist(exam_data$MATHS)
```

default setting of bins are 30

```{r}
ggplot(data=exam_data,aes(x=MATHS))+
  geom_histogram(bins=10, 
                color='red',
                 fill='blue')+
  ggtitle("Histogram")
```

```{r}
ggplot(data=exam_data,aes(x=MATHS,fill=GENDER))+
  geom_histogram(bins=20, color="grey30")#color in histogram is bin line color
```

## Density plot

it is a good way to replace histogram for continuous data if its distribution is smooth

```{r}
ggplot(data=exam_data,aes(x=MATHS,colour=GENDER))+
  geom_density() 
```

## Dotplot

```{r}
ggplot(data=exam_data,aes(x=MATHS))+
  geom_dotplot(dotsize=0.5,binwidth = 3)+
  scale_y_continuous(NULL,breaks = NULL) #remove y-axis value
```

## Boxplot and Notch

Notch(缺口) can check by whether notches are overlapping to see whether medians are the same

```{r}
ggplot(data=exam_data,aes(x=GENDER,y=MATHS))+
  geom_boxplot(notch=TRUE)
```

## Violin Plot

Can compare more clearly about the difference of distributions of multiple data

```{r}
ggplot(data=exam_data,aes(x=GENDER,y=MATHS))+
  geom_violin()
```

## Scatter plot

```{r}
ggplot(data=exam_data,aes(x=ENGLISH,y=MATHS))+
  geom_point()+
  geom_smooth()+
  coord_cartesian(xlim=c(0,100),ylim=c(0,100))
```

### Combine scatter and boxplot

```{r}
ggplot(data=exam_data,aes(x=GENDER,y=MATHS))+
  geom_point(position="jitter",size=0.2)+
  geom_boxplot()
```

### Statistical summary

Add statistical summary can use stat_summary or overwrite by a new geom\_ setting

```{r}
ggplot(data=exam_data,aes(x=GENDER,y=MATHS))+
  geom_point(position='jitter')+
  geom_boxplot()+
  stat_summary(geom="point",fun='mean',color='red',size=4)
```

```{r}
ggplot(data=exam_data,aes(x=GENDER,y=MATHS))+
  geom_boxplot()+
  geom_point(stat="summary",        
             colour ="red",          
             size=4)
```

## Create 2D panel

```{r}
ggplot(data=exam_data,aes(x=MATHS))+
  geom_histogram(bins=20)+
  facet_wrap(~CLASS)
```

### Facet_grid

Two discrete variables can use facet_grid

```{r}
ggplot(data=exam_data,aes(x=MATHS))+
  geom_histogram(bins=20)+
  facet_grid(~CLASS)
```
