---
title: "Hands-on_Ex03"
---

## Import package

```{r,include=FALSE}
#| echo: false
pacman::p_load(ggiraph, plotly, 
               patchwork, DT, tidyverse, readxl, gifski, gapminder,
               gganimate, tidyverse)
```

# Interactive Data Visualisation

## Import data

```{r}
exam_data <- read.csv("data/Exam_data.csv")
```

## Interactive graph with ggiraph,tooltip

### Tooltip effect with tooltip aesthetic

1.  Tooltip: show label when mouse scroll on it
2.  histodot: dot stacks vertically
3.  svg: Scalable Vector Graphics(adjust size so can with good resolution) for html page
4.  ratio of 0.618 is a recommended ratio for visually pleasing

```{r}
plot <- ggplot(data=exam_data,aes(x=MATHS))+
  geom_dotplot_interactive(aes(tooltip=ID),
                           stackgroups=TRUE,
                           binwidth=1,
                           method='histodot')+
    scale_y_continuous(NULL,breaks=NULL)
girafe(
  ggobj=plot,
  width_svg = 6,
  height_svg = 6*0.618
)
```

### Displaying multiple information

1.  Use "\$" to extract specific column from dataframe
2.  "paste0" to concatenate string

```{r}
exam_data$tooltip <- c(paste0(
  "Name =", exam_data$ID, "\n Class =", exam_data$CLASS
))
plot <- ggplot(data=exam_data,aes(x=MATHS))+
  geom_dotplot_interactive(aes(tooltip=exam_data$tooltip),
                           stackgroups=TRUE,
                           binwidth=1,
                           method='histodot')+
    scale_y_continuous(NULL,breaks=NULL)

girafe(
  ggobj=plot,
  width_svg = 8,
  height_svg = 8*0.618
)

```

### Customising Tooltip

Use opts_tooltip() to customize interactive label s font and color

```{r}
tooltip_css <- "background-color:white;
font-style:bold; color:blue;"

plot <- ggplot(data=exam_data,aes(x=MATHS))+
  geom_dotplot_interactive(aes(tooltip=ID),
                           stackgroups=TRUE,
                           binwidth=1,
                           method="histodot")+
  scale_y_continuous(NULL,breaks=NULL)

girafe(
  ggobj=plot,
  width_svg = 6,
  height_svg = 6*0.618,
  options = list(
    opts_tooltip(
      css=tooltip_css))
)

```

### Displaying statistics

1.  accuracy: 0.01 means to round to two decimal places

2.  ymax: upper limit of 90% interval of y, y: average

3.  scales::number is to format value with accuracy (0.01)

4.  after_stat:tooltip applied after calculating statistic values

5.  fun.data: summary statistics, mean_se:calculate mean and standard error

```{r}
tooltip <- function(y,ymax,accuracy=0.001){
  mean <- scales::number(y,accuracy=accuracy)
  se <- scales::number(ymax-y,accuracy = accuracy)
  paste0("Mean math scores",mean," +/- ",se)
}

gg_point <- ggplot(data=exam_data,aes(x=RACE)
               )+
  stat_summary(aes(y=MATHS,
                   tooltip=after_stat(
                     tooltip(y,ymax))),
  fun.data="mean_se",
  geom=GeomInteractiveCol,
  fill="lightblue"
    )+
  stat_summary(aes(y=MATHS),
  fun.data="mean_se",
  geom="errorbar",width=0.2,size=0.2)

girafe(
  ggobj=gg_point,
  width_svg = 8,
  height_svg = 8*0.6188)
```

### Highlight effect

data_id allows highlighting effect on dotplot with defined category, default setting is filling with orange color.

```{r}
plot <- ggplot(data=exam_data,aes(x=MATHS))+
  geom_dotplot_interactive(aes(data_id=CLASS),
                           stackgroups=TRUE,
                           binwidth=1,
                           method="histodot")+
  scale_y_continuous(NULL,breaks=NULL)

girafe(
  ggobj=plot,
  width_svg = 6,
  height_svg=6*0.618
)
```

### Changing hovering color setting

opacity: changing the transparency, opacity=0.2 means with 80% transparency

```{r}
plot <- ggplot(data=exam_data,aes(x=MATHS))+
  geom_dotplot_interactive(aes(data_id=CLASS),
                           stackgroups=TRUE,
                           binwidth=1,
                           method="histodot")+
  scale_y_continuous(NULL,breaks=NULL)

girafe(
  ggobj=plot,
  width_svg = 6,
  height_svg=6*0.618,
  options = list(
    opts_hover(css="fill: #202020;"),
    opts_hover_inv(css="opacity:0.1;")
    
  )
)
```

### Combine tooltip(label) and hovering effect

```{r}
plot <- ggplot(data=exam_data,aes(x=MATHS))+
  geom_dotplot_interactive(
    aes(tooltip=CLASS,data_id=CLASS),
    stackgroups=TRUE,
    binwidth=2,
    method="histodot"
  )+
  scale_y_continuous(NULL,breaks=NULL)

girafe(
  ggobj=plot,
  width_svg=6,
  height_svg=6*0.618,
  options = list(
    opts_hover(css="fill:#202020;"),
    opts_hover_inv(css="opacity:0.2;")
  )
  
)
```

### Insert link on the graph

1.  add onclick column to exam_data
2.  sprintf: combine new string with URL
3.  "%s%s", first %s is the link, second %s is new adding string
4.  as.character: convert to character(string) type, and adding to URL

```{r}
exam_data$onclick <- sprintf("window.open(\"%s%s\")",
 "https://www.moe.gov.sg/schoolfinder?journey=Primary%20school",
 as.character(exam_data$ID))

plot <- ggplot(data=exam_data,aes(x=MATHS))+
  geom_dotplot_interactive(
    aes(onclick=onclick),
    stackgroups=TRUE,
    binwidth=2,
    method="histodot"
  )+
  scale_y_continuous(NULL,breaks=NULL)

girafe(
  ggobj=plot,
  width_svg=6,
  height_svg=6*0.618
  
)
 
 
```

### Coordinate multiple views

1.  opts_hover: setting about hovering point
2.  opts_hover_inv: setting for other part when hovering

```{r}

plot1 <- ggplot(data=exam_data,aes(x=MATHS))+
  geom_dotplot_interactive(
    aes(data_id=ID),
    stackgroups=TRUE,
    binwidth=2,
    method="histodot"
  )+
  coord_cartesian(xlim=c(0,100))+
  scale_y_continuous(NULL,breaks=NULL)

plot2 <- ggplot(data=exam_data,aes(x=ENGLISH))+
  geom_dotplot_interactive(
    aes(data_id=ID),
    stackgroups=TRUE,
    binwidth=2,
    method="histodot"
  )+
  coord_cartesian(xlim=c(0,100))+
  scale_y_continuous(NULL,breaks=NULL)

girafe(code = print(plot1+plot2),
  width_svg=6,
  height_svg=3,
  options=list(
    opts_hover(css="fill: #202020;"),
    opts_hover_inv(css="opacity:0.2;")
  
))
 
```

## Interactive graphs by plotly

### Interactive scatter plot

```{r}
#|message: false
plot_ly(data = exam_data, 
             x = ~MATHS, 
             y = ~ENGLISH)
```

### Adding visual effect

```{r}
#| warning: false
plot_ly(data= exam_data,
       x= ~MATHS,
       y= ~ENGLISH,
       color= ~RACE)
```

### Coordinate multiple views

1.  highlight_key: create shared data for two plots
2.  subplot: coordinate two plots together

```{r}
d <- highlight_key(exam_data)

p1 <- ggplot(data=d,aes(x=MATHS,y=ENGLISH))+
  geom_point(size=2)+
  coord_cartesian(xlim=c(0,100),ylim=c(0,100))

p2 <- ggplot(data=d,aes(x=MATHS,y=SCIENCE))+
  geom_point(size=2)+
  coord_cartesian(xlim=c(0,100),ylim=c(0,100))

subplot(ggplotly(p1),ggplotly(p2))
```

## Interactive graphs by ggplotly

```{r}
plot <- ggplot(data=exam_data,aes(x=MATHS,y=ENGLISH))+
  geom_point(size=2)+
  coord_cartesian(xlim=c(0,100),ylim=c(0,100))


ggplotly(plot)
```

### Data table

class="compact" means to have compact layout for the table

```{r}
DT::datatable(exam_data,class="compact")
```

## Link table with graph

1.  highlight_key: assign identifier to dataframe

2.  plotly_selected: track corresponding point when mouse click on table row

3.  crosstalk: combine table and graph

4.  bscols: create layout with two elements

```{r}
data <- highlight_key(exam_data)

p <- ggplot(data,aes(ENGLISH,MATHS))+
  geom_point(siz=1)+
  coord_cartesian(xlim=c(0,100),ylim=c(0,100))

gg <- highlight(ggplotly(p),"plotly_selected")

crosstalk::bscols(gg,DT::datatable(data),widths = 5)
```

# Animated Statistical Graphics

## Import data
1. mutate_each or mutate(across())：apply given function to selected columns
2. %>%  pass function/results to next step
3. funs all all_of: create a list of function
4. factor: turn column into categorical variable
```{r}
col <- c("Country","Continent")
global <- read_xls("data/GlobalPopulation.xls",
                   sheet="Data") %>% 
  mutate(across(all_of(col), factor)) %>% 
  mutate(Year=as.integer(Year))

```

## Population bubble plot
### static plot

```{r}
ggplot(global,aes(x=Old,y=Young,size=Population,colour=Country))+
  geom_point(alpha=0.7,show.legend = FALSE)+
  scale_color_manual(values=country_colors)+
  scale_size(range=c(2,10))+
  labs(title='Year: {frame_time}', 
       x = '% Aged', 
       y = '% Young')
```
### Animated plot with gganimate
ease_aes(): the moving way of plot element, default is linear
```{r}
ggplot(global,aes(x=Old,y=Young,size=Population,colour=Country))+
  geom_point(alpha=0.7,show.legend = FALSE)+
  scale_color_manual(values=country_colors)+
  scale_size(range=c(2,10))+
  labs(title='Year: {frame_time}', 
       x = '% Aged', 
       y = '% Young')+
  transition_time(Year)+
  ease_aes("linear")
```

### Animated plot with ggplotly

```{r}
plot <- ggplot(global,aes(x=Old,y=Young,size=Population,colour=Country))+
  geom_point(aes(size=Population, frame=Year),
             alpha=0.7,show.legend = FALSE)+
  scale_color_manual(values=country_colors)+
  scale_size(range=c(2,10))+
  labs( x = '% Aged', 
       y = '% Young')

ggplotly(plot)
```

### Using plot_ly()

```{r,warning=FALSE}
bp <- global %>%
  plot_ly(x = ~Old, 
          y = ~Young, 
          size = ~Population, 
          color = ~Continent, 
          frame = ~Year, 
          text = ~Country, 
          hoverinfo = "text",
          type = 'scatter',
          mode = 'markers'
          )
bp
```

