---
title: "Hands-on Exercise 2"
author: "Yen Yun Hsuan"
---

```{r,message=FALSE,include=FALSE}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse)
library(ggplot2)
library(extrafont) 
```

```{r}
exam_data <- read.csv("data/Exam_data.csv")
```

### ggrepel

An extension of ggplot that can deal with overlapping text For example, add "\_repel" to geom_label

```{r}
ggplot(data=exam_data,aes(x=MATHS,y=ENGLISH))+
  geom_point()+
  geom_smooth(method=lm,linewidth=0.5)+
  geom_label_repel(aes(label=ID),fontface='bold')+ 
  coord_cartesian(xlim=c(0,100),ylim=c(0,100))+ 
  ggtitle('English scores versus Maths scores for Primary 3"')
  
```

### ggplot theme

```{r}
p1 <- ggplot(data=exam_data,aes(x=MATHS))+
  geom_histogram(bins=20,
                 boundary=100,
                 color='grey25',
                 fill='grey90')+
  theme_gray()+
  ggtitle("Distribution of math scores")
p1
```

```{r}
ggplot(data=exam_data,aes(x=MATHS))+
  geom_histogram(bins=20,
                 boundary=100,
                 color='grey25',
                 fill='grey90')+
  theme_economist()+
  ggtitle("Distribution of math scores")
```

### Working with hrbthems package

More fonts and label placements choices, adjust by axis_title size, base_size, axis_text_size

```{r,warning=FALSE}
ggplot(data=exam_data,aes(x=MATHS))+
  geom_histogram(bins=20,boundary=100,
                 fill='grey25',
                 color='grey90')+
  ggtitle("Distribution of math scores")+
  theme_ipsum(axis_title_size = 20,base_size=20,grid='Y') 

```

### Beyond Single Graph

```{r}
p2 <- ggplot(data=exam_data,aes(x=ENGLISH))+
  geom_histogram(bins=20, boundary=100,fill='grey20',color='grey90')+
  ggtitle("Distribution of English score")+
  coord_cartesian(xlim=c(0,100))
p2
```

```{r,warning=FALSE}
p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
p3
```

#### Combining two ggplot2 graphs

```{r,warning=FALSE}
p1|p2
```

"\|" operator to stack two ggplot2 graphs,

"/" operator to place the plots beside each other,

"()" operator the define the sequence of the plotting

#### Add tag

```{r,warning=FALSE}
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```

#### Place plots on top of below other plots

```{r,warning=FALSE}
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```

#### Use patchwork and ggtheme

```{r}
patchwork <- p3| (p1 / p2)
patchwork & theme_economist()+theme(plot.title = element_text(size = 9))

```
